Assessing the Replicability of Psychological Science Through Pedagogy

### 2016-06-24 paper/paper.Rmd
methods and intro

author-information:
    - \author{Eric Smith*, Robert X. D. Hawkins*, Michael C. Frank}

@@

With respect to broader questions of meta-science, it is a challenging proposition to estimate the reproducibility of an entire literature [cf. @osc2015;@gilbert2016;@anderson2016].

In sum, we view the contribution of the current work to be two-fold. First, we provide a proof-of-concept and pedagogical model for conducting replications in the classroom. Second, we provide an estimate of how likely it is for a graduate student to choose an article of interest in *Psychological Science* and -- within constraints of budget, expertise, and effort -- reproduce the basic pattern of findings with sufficient fielility to be able to build on it in future work. We believe that this notion of whether a piece of research supports cumulative progress is an important construct. Whether another independent scientist can build on a result is perhaps as important as -- and in many cases easier to measure -- whether the work is reproducible in some more abstract sense.

@@

names(ppt) = c("Citation", "Expt Num", "Data", "Materials", "Turk?", 
               "N (orig)","N (rep)")

@@

% latex table generated in R 3.3.0 by xtable 1.8-2 package
% Fri Jun 24 12:02:11 2016
\begin{table}[ht]
\centering
\begin{tabular}{lllllrr}
  \hline
Citation & Expt Num & Data & Materials & Turk? & N (orig) & N (rep) \\ 
  \hline
@storm2015 & 3 & No & No & No &  48 &  61 \\ 
@lewis2015 & 4 & No & Yes & Yes & 122 & 128 \\ 
@scopelliti2015 & 3 & No & Yes & Yes & 550 & 124 \\ 
@liverence2015 & 1 & No & Some & No &  18 &  \\ 
@wang2015 & 2 & No & No & No & 219 & 397 \\ 
@sofer2015 & 1 & No & Yes & No &  48 &  95 \\ 
@ko2015 & 2 & Yes & Some & No &  40 &  40 \\ 
@dunning2015 & 1b & No & Yes & Yes & 202 &  50 \\ 
@proudfoot2015 & 1 & No & No & Yes &  80 &  84 \\ 
@zaval2015 & 1 & Yes & Yes & Yes & 312 & 328 \\ 
@franconeri2015 & 1a & No & No & No &  12 &  27 \\ 
   \hline
\end{tabular}
\caption{Summary of attempted replications} 
\label{tab:summary}
\end{table}

@@

Once requested changes were made, the student conducted "Pilot B," using a handful of naive participants recruited from Mechanical Turk. The goal of this pilot was to ensure that participants were able to complete the experiment and did not report any substantial issues with instructions or technical details of the experiment. (All students were instructed to give participants a way to leave free-form comments at the end of the experiment but prior to debriefing text). At the conclusion of Pilot B, both the instructor and a TA reviewed the student's analytic code and its outputs on the data for all confirmatory analysis. The goal of this code review was to ensure that all planned analyses were

Contributions
Pedagogical recipe
Author contact template
Course experiment templates
Extraction of key statistic, power analysis
Multiple reviews by team
Pre-registration
Proof of concept
Table w/ all included replications
Scatter plot of effect sizes?
Actual findings
## Individual Replications

Details of individual replications are available in the supplementary materials.  
### @storm2015

### @lewis2015

### @scopelliti2015

### @liverence2015

### @wang2015

### @ko2015

### @dunning2015

@@

Analyses for the final project were pre-registered ([https://osf.io/rxz8m/]()).

>>>>

author-information:
    - \author{Eric Smith*, Robert X. D. Hawkins*, the students of Psych 254, & Michael C. Frank}

@@

With respect to broader questions of meta-science, it is a challenging proposition to estimate the replicability of an entire literature [cf. @osc2015;@gilbert2016;@anderson2016]. Making such an estimate requires determining a sampling strategy, a notion of what constitutes a direct replication, what constitutes a successful replication, among many other difficult decisions. In the current work, we focus on a quantity that is easier to estimate: the probability that a graduate student can choose an article of interest in *Psychological Science* and -- in a single attempt, and within constraints of budget, expertise, and effort -- reproduce the basic pattern of findings with sufficient fidelity to be able to build on it in future work. We believe that this notion of whether a piece of research supports cumulative progress is an important construct. Whether another independent scientist can build on a result is perhaps as important as -- and in many cases easier to measure -- whether the work is reproducible in some more abstract sense.

In sum, we view the contribution of the current work to be two-fold. First, we provide a proof-of-concept and pedagogical model for conducting replications in the classroom. Second, we provide an initial estimate of what proportion of studies from the most recent volume of *Psychological Science* could easily be reproduced by a motivated student.

@@

names(ppt) = c("Citation", "Expt", "Data", "Materials", "Turk?", 
               "N (orig)","N (rep)")

@@

% latex table generated in R 3.3.0 by xtable 1.8-2 package
% Fri Jun 24 12:02:11 2016
\begin{table}[ht]
\centering
\begin{tabular}{lllllrr}
  \hline
Citation & Expt & Data & Materials & Turk? & N (orig) & N (rep) \\ 
  \hline
@storm2015 & 3 & No & No & No &  48 &  61 \\ 
@lewis2015 & 4 & No & Yes & Yes & 122 & 128 \\ 
@scopelliti2015 & 3 & No & Yes & Yes & 550 & 124 \\ 
@liverence2015 & 1 & No & Some & No &  18 &  \\ 
@wang2015 & 2 & No & No & No & 219 & 397 \\ 
@sofer2015 & 1 & No & Yes & No &  48 &  95 \\ 
@ko2015 & 2 & Yes & Some & No &  40 &  40 \\ 
@dunning2015 & 1b & No & Yes & Yes & 202 &  50 \\ 
@proudfoot2015 & 1 & No & No & Yes &  80 &  84 \\ 
@zaval2015 & 1 & Yes & Yes & Yes & 312 & 328 \\ 
@franconeri2015 & 1a & No & No & No &  12 &  27 \\ 
   \hline
\end{tabular}
\caption{Summary of attempted replications} 
\label{tab:summary}
\end{table}

@@

Once requested changes were made, the student conducted "Pilot B," using a handful of naive participants recruited from Mechanical Turk. The goal of this pilot was to ensure that participants were able to complete the experiment and did not report any substantial issues with instructions or technical details of the experiment. (All students were instructed to give participants a way to leave free-form comments at the end of the experiment but prior to debriefing text). At the conclusion of Pilot B, both the instructor and a TA reviewed the student's analytic code and its outputs on the data for all confirmatory analysis. The goal of this code review was to ensure that all planned analyses were specified correctly and with sufficient detail to permit inclusion in the broader analysis.

After Pilot B review was completed, students were given authorization to collect the full sample of participants, contingent on having made the requested change. Prior to data collection, the analytic script containing all confirmatory analyses was pre-registered using the Open Science Framework. (In one case, a student failed to complete the pre-registration but nevertheless submitted a pre-specified analysis plan to the instructors for review).

## Sample size determination and statistical approach

After selecting a paper of interest, students selected a particular study from within that paper based on their judgment of interest and feasibility. The modal study chosen was the first, but several students chose later studies in a set as well.

For purposes of planning and analysis, we asked stuents Despite the general emphasis in meta-analytic and reproducibility efforts to aggregate a single effect size of interest [@osc2015;@lipsey2001], it was often challenging to identify a single key statistical test in studies. Often a wide variety of statistical tests were conducted, often within one or several regression models with multiple specifications. Thus, we interpret "key statistical test" results with caution. Nevertheless, despite this interpretive difficulty, we attempted to use a single statistical test as our target for experiment planning and analysis. This test was determined by iterative discussion between the student, instructor, and TAs.

Derermining appropriate sample sizes is a major challenge in replication research [@button2013,@simonsohn2015]. Post-hoc power analyses are problematic because of the likelihood of inflation of effect sizes due to the "winner's curse" [@hoenig2001;button2013], but -- especially with the constraints of a limited budget -- it may be impossible to follow more conservative guidelines in all cases. For example, @simonsohn2015 recommends 2.5x the original sample, which can be feasible for small or underpowered original studies but may lead to impractical or unnecessary recommendations for experiments that were initially large and/or adequately powered.

We eventually used a hybrid, case-by-case selection criterion that attempted to maximize the success of the project while staying within the constraints of our budget.^[We initially were allocated $1,500 but one student contributed personal research funds, leading to a total cost of \approx $1,700 for all studies. We attempted to set payment for our studies at approximately $6/hour based on timing estimated from pilot A sessions.] In `{r} sum(project_plan$power_standard == "80% power")} cases, we powered the replication attempt to 80% post-hoc power. In `{r} sum(project_plan$power_standard == "Original")` cases, we used the original sample size. In one case, the original experiment was apparently extremely over-powered in the post-hoc analysis, so we chose a smaller sample size that still would apparently yield high power.  In one case, we reduced the number of trials as part of the adaptation to mechanical turk, and in one case of a multi-step procedure where the first step consisted of stimulus generation, we decreased the N in the first stage of the procedure (details available in Supplemntal Materials).

## Individual Replications

Details of individual replications are available in the supplementary materials. Students chose findings from a wide variety of domains for replication. One group was nvestigations of memory and visual attention [e.g., @franconeri2015], including in applied contexts like remembering "deleted" files [@storm2015] or tracking transition between locations on smartphone apps [@liverence2015]. A second cluster came from studies of social perception, including judgments of faces and voices [@ko2015;@sofer2015], as well as studies of attributions of modesty, creativity, and expertise [@atir2015;@scopelliti2015;@proudfoot2015]. Other studies investigated increasing retirement savings by orienting participants to the future [@lewis2015], legacy priming [@zaval2015], and the effects of math anxiety on performance [@wang2015].

## Analysis strategy

Analyses for the final project were pre-registered ([https://osf.io/rxz8m/]())

@@

.
